The pollination trade-off

### 2018-07-20 paper/manuscript.Rmd
Edit intro to add motivation about traits, pollen and abundance

## Introduction

@@

This is so, because sharing a mutualistic partner also has costs. 
The end result is that, when the cummulative costs have negative impact on fitnessis, species also start competeing for mutualistic partners. 
For the 88% of flowering plants that rely on animal pollination, there are two mechanisms trough which competition for pollinators might arrise. 
The first is competition for visits, where plants with more atractive flowers reduce the pollination service for those less atractive. 
The second is interspecific pollen transfer, where even receiving a visit might not translate into net facilitation. 
A focal plant might receive pollen from a different species, or conversely pollen from the focal plan might be lost to different species. 
Here we focus on this second mechanism.

The potential negative effects of competition for pollinators have been long recognised and widely documented. 
It is currently clear that they have the potential to drive the evolution of flower phenologies, morphologies and reproductive strategies as well as driving ecological diferentiation. 
Indeed there is considerable evidence demonstrating the effects that interspecific pollen transfer in two-species systems. 
However, these competitive interactions between species pairs do not occur in isolation. 
Instead, plants often ocurr in communities in which multiple competitive interactions operate simultaneusly and lead to emergent phenomena not observed at smaller scales.

Nevertheless, only a handful of studies have examined interspecific pollen transfer at the community level, and therefore we still understand little about how compeition for pollinators operates at this scale. 
All together, existing studies show that, as expected, the effective result of mutualistic interactions is, in most cases, one of facilitation. 
Importantly, they also show a large variability in this result—both across species and across communities. 
However, the driving factors behind this variability are currently not understood.

Understanding what makes species and communities different is important not only because of its ecological and evolutionary implications. 
It is also important because it is fundamentaly related to the way we understand and mathematically represent mutualistic systems. 
Some of the most widely used theoretical models of pollination account for the negative effects of mutualisim by adjusting the mutualistic benefit of an interaction so that it can be negatively affected by the number of mutualistic partners---the species degree. 
The extent to which species degree affects the benefit is called *mutualistic trade-off*. 
This approach, considerably simplifies the mathematical treatment of competition for pollinators. 
At the same time it is based on the intuitive rationale that a species that interacts with few species benefits strongly from each of them, whereas a species that interacts with a large number of species does so comparatively weakly. 
Although, the mutualistic trade-off has been shown to play a determining role on the stability and diversity of mutualistic communities, empirical evidence is scarce, to say the least.

Elucidating how the mechanisms and drivers of facilitation scales to the community level and its theoretical concequences has been elusive in part because of the large amounts of data required.
Previous studies focus on the final picture of interspecific pollen transfer that is obtained by examining the pollen deposited on flower stigmas. 
However, teasing appart the competition for pollinators also requires data that describes the structure of the interactions and the pollen flows that ocurr within it, while accounting for the traits of the species that integrate the community.
Here, we examine a comprehensive data set that includes information about pollen deposition, transfer, visitation, species abundance, and species traits across diverse communities in the Argentinean Pampas. 
Specifically, we first estimate the gain on pollen that can be attributed to the mutualistic interactions. 
Second, we investigate wether there is a relationship between this gain and a suite of variables that describe both the role of the plant species and the ecological context. 
Here, we pay special attention to the role of species degree because of its ramifications to theoretical ecology. 
At the same time, we examine wether the plant role and ecological context is best infered from dynamics occurring at the community level, or wether it can be well approximated using information available from all sites in which a species ocurr. 
Third, and finaly, we explore how plant species differ on the extent to which competition for pollinators shapes facilitation.

## Methods

>>>>

# Introduction

@@

This is so, because sharing a mutualistic partner also has costs. 
The end result is that, when the cummulative costs have a negative impact on fitness, species also start competing for mutualistic partners. 
The possible negative effects of competition for pollinators have been long recognised and widely documented.
Based mainly on two-species systems, it is currently clear that they have the potential to drive the evolution of flower phenologies, morphologies and reproductive strategies as well as driving ecological diferentiation. 
However, these competitive interactions between species pairs do not occur in isolation. 
Plants often ocurr in communities in which multiple competitive interactions operate simultaneusly and lead to emergent phenomena not observed at smaller scales. 
Here, we explore the signatures competition for pollinators in diverse natural communities, where empirical data is more scarce.

At the community scale, competition for pollination might arrise from changes in the quantity and quality of pollination. 
Quantity is mainly affected by competition for visits, where plants with more atractive flowers reduce the number of visits for those less atractive and therefore they may experience a reduction in the amount of conspecific pollen received. 
Quality is mainly affected by a more subtle but perhaps more important mechanism—interspecific pollen transfer, where even receiving a visit might not translate into net facilitation. 
This is so because focal plant might receive pollen from a different species, or conversely pollen from the focal plant might be lost to different species. 
Generally speaking the larger the proportion of conspecific relative to heterospecific pollen the higher the quality of the pollination service.

Only a handful of studies have examined interspecific pollen transfer at the community level, and therefore we still understand little about how compeition for pollinators operates at this scale. 
All together, existing studies at the community scale show that, as expected, the effective result of mutualistic interactions is, in most cases, one of facilitation. 
Importantly, they also show a large variability in this result—both across species and across communities. 
However, the driving factors behind this variability are currently not understood.

One of the most intuitive drivers of this variability, and therefore widely used in theoretical models of mutualism to account for the neggative effects of mutualism is the species degree---the number of mutualistic partners. 
It is based on the intuitive rationale that a species that interacts with few species benefits strongly from each of them, whereas a species that interacts with a large number of species does so comparatively weakly. 
Recent empirical evidence suggest that other factors might also play an important role shaping the competition for pollination. 
For instance, species with original traits (far from the community centroid) have been shown to generally have fewer interaction partners. 
If evolutionary specialisation occur by changing traits to focus on fewer but better partners, we should expect a reduction of competition for pollination in species with original traits. 
Similarly, it might also be the case that density is the dominant force driving pollen transfer. 
Abundant species might experience a dilution of the available pollinators but might also comprise a greater component of the pollen source pool being transported by pollinators.  
In this case a potential reduction in the quantity of pollination could be compensated by an increase in the quality of pollination.

Elucidating how the mechanisms and drivers of facilitation scales to the community level and its theoretical concequences has been elusive in part because of the large amounts of data required.
Previous studies focus on the final picture of interspecific pollen transfer that is obtained by examining the pollen deposited on flower stigmas. 
However, as we have described, teasing appart competition for pollinators also requires data that describes the structure of the interactions and the pollen flows that ocurr within it, while accounting for the traits of the species that integrate the community. 
Here, we examine a comprehensive data set that includes information about pollen deposition, transfer, visitation, species abundance, and species traits across diverse communities in the Argentinean Pampas.

## Methods
First, we estimate the gain on pollen that can be attributed to the mutualistic interactions. 
Second, we investigate wether there is a relationship between this gain and a a suite of variables that together describe the role of the plant species and the ecological context. 
Third, and finaly, we explore how these variables shape the quantity and quality of pollination and ultimately the competition for pollinators.
**TODO: Mini conclusion here**
